### 2021-07-29 thesis/tex/ICtheory.tex
update

\section{IceCube}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact with charged lepton in the ice. We now are interested in the effective area $\Aeff$, 
i.e.~the cross-section of the detector that the lepton is exposed to.
$\Aeff$ depends on several parameters, some of them being detector physical volume, $\Etrue$,$\ztrue$ and the neutrino cross-section. Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
The data file has the following form

@@

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$.
Just as with the fluxes, we interpolate this in $\Etrue,\ztrue$ and show the result %TODO aeff interpolation

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}}\label{fig:IC_data}
\end{figure}

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more persicely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `offical', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor which contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source what will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes atmospheric prompt and astrophysical fluxes. The IceCube collaboration use several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}

In Fig.~\ref{fig:IC_MC_norm}, we present the Icecube Monte Carlo obtained from their 2020 sterile analysis~\cite{IC2020}, along
with our null hypothesis times a constant factor.  We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}. We now arrive to our final event count
\begin{align}\label{eq:Nth}
    N_{ij} &= \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} T \sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_{E^t_{min}}^{E^t_{max}} R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta\, \dd E^t
    \,,
\end{align}
with $E^t_{min}$, $E^t_{max}$, and $R(E^r,E^t)$ taken from the Gaussian process regressor.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as ´Null' in the plots.}\label{fig:IC_MC_norm}
\end{figure} 
Thus, we are now able to sufficiently approximate the IceCube Monte Carlo, which makes us able to run simulations based on different physics scenarios.
% \newpage
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

>>>>

\section{IceCube}\label{ch:ICmethod}
As the neutrinos have propagated the Earth, they arrive at the South Pole, where they interact with charged lepton in the ice. We now are interested in the effective area $\Aeff$, 
i.e.~the cross-section of the detector that the lepton is exposed to.
$\Aeff$ depends on several parameters, some of them being detector physical volume, $\Etrue$, $\ztrue$ and the neutrino cross-section. Fortunately, the binned $\Aeff$ is provided to us by the collaboration~\cite{ICaeff}.
The data file has the following form

@@

Here, $\Aeff$ has been averaged over $\Aeff_\mu$ and $\Aeff_{\bar{\mu}}$.
Just as with the fluxes, we interpolate this in $\Etrue$, $\ztrue$ and show the result %TODO aeff interpolation

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}.}\label{fig:IC_data}
\end{figure}

\subsection*{Monte Carlo normalization}
Independent researchers outside of the IceCube collaboration will not be able to more percisely
simulate the detector. The IceCube Monte Carlo is a complex and proprietary machinery, so our goal in this 
section is to come as close as we can to their Monte Carlo simulations. After we are confident that 
our code displays the same overall features as the `offical', we normalize our results $N_{ij}^\text{sim}$ as 
\begin{align}\label{eq:MC_norm}
    N_{ij} = \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} N_{ij}^\text{sim}\,.
\end{align}
For each bin $i,j$, we then obtain a correction factor which contains information that we are unable
to obtain or sufficiently incorporate. One example of such information is the systematic errors of the DOMs.
Recent IceCube data releases do not include such information. Since the systematic errors are affecting the 
event count on a bin-by-bin basis, they can in theory drastically modify the binned results. Another example of
an error source what will be remedied by this method is the flux. We are using a fairly simple model of the atmospheric 
flux that excludes atmospheric prompt and astrophysical fluxes. The IceCube collaboration use several different flux models which are initialized 
by a parametrization of the cosmic ray flux.\footnote{Included in the cosmic ray models are e.g. the pion to kaon 
ratio, which are often used as a nuisance parameter. By not being able to include this in our error analysis, our method will 
be limited to only consider the overall flux normalization, rather than the components that produce the flux in the first place.}

In Fig.~\ref{fig:IC_MC_norm}, we present the Icecube Monte Carlo obtained from their 2020 sterile analysis~\cite{IC2020}, along
with our null hypothesis times a constant factor.  We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}. We now arrive to our final event count
\begin{align}\label{eq:Nth}
    N_{ij} &= \frac{N_{ij}^\text{null}}{N_{ij}^\text{MC}} T \sum_\beta \int_{(\cos{\theta_z^r})_i}^{(\cos{\theta_z^r})_{i+1}} \dd \cos{\theta^r_z} \int_{E^r_{j}}^{E^r_{j+1}} \dd E^r 
    \int_{E^t_{min}}^{E^t_{max}} R(E^r,E^t) \phi_\beta^\text{det}  A^\text{eff}_\beta\, \dd E^t
    \,,
\end{align}
with $E^t_{min}$, $E^t_{max}$, and $R(E^r,E^t)$ taken from the Gaussian process regressor.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as `Null' in the plots.}\label{fig:IC_MC_norm}
\end{figure} 
Thus, we are now able to sufficiently approximate the IceCube Monte Carlo, which makes us able to run simulations based on different physics scenarios.
% \newpage
% \bibliographystyle{nature}
% \bibliography{ref.bib}
% \end{document}

### 2021-07-29 thesis/tex/nsiTheory.tex
update

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b},\, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as the open up matter interactions across flavors. Remember, in the Standard model,
we are restriced to weak interactions that conserve lepton number. However, a off-diagonal NSI parameter allows flavor transitions during matter interactions.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition which is supressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.

>>>>

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b},\, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as they open up matter interactions across flavors. Remember, in the Standard model,
we are restriced to weak interactions that conserve lepton number. However, a off-diagonal NSI parameter allows flavor transitions during matter interactions. Thus,
the off-diagonal elements constitute new soures of flavor violation.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition which is supressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.

### 2021-07-29 thesis/tex/sterileTheory.tex
update

\subsection{Sterile Signals}
Since the new neutrino does not interact weakly, how do we then detect its signal? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappeareance in the for $P_{\anm\anm}$, shown in Fig.~\ref{fig:sterile_resonance}.

>>>>

\subsection{Sterile Signals}
Since the new neutrino does not interact weakly, how do we then detect its signal? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappeareance in the for $P_{\bar{\mu}\bar{\mu}}$, shown in Fig.~\ref{fig:sterile_resonance}.
